---
title: "Clustering Lab"
author: "Sam Muir"
date: "2024-02-29"
output: html_document
---

```{r, echo = FALSE, eval = TRUE, message=FALSE}
library(tidyverse) 
library(cluster) #cluster analysis
library(factoextra) #cluster visualization
library(tidymodels) #simulation 
library(readr) #read data
library(RColorBrewer)# Color palettes

```

We'll start off with some simulated data that has a structure that is amenable to clustering analysis.

```{r init_sim}
#Set the parameters of our simulated data
set.seed(101)

cents <- tibble(cluster = factor(1:3),
                num_points = c(100,150,50),
                x1 = c(5,0,-3),
                x2 = c(-1,1,-2))
```


```{r sim}
#Simulate the data by passing n and mean to rnorm using map2()

labelled_pts <- cents %>%
  mutate(x1 = map2(num_points, x1, rnorm),
         x2 = map2(num_points, x2, rnorm)) %>%
  select(-num_points)  %>%
  unnest(cols = c(x1,x2))

ggplot(labelled_pts, aes(x1, x2, color = cluster)) +
  geom_point(alpha = 0.5)
```


```{r kmeans}
points <- labelled_pts %>%
  select(-cluster)
  
kclust <- kmeans(points, centers = 3, n = 1)
kclust
```

```{r syst_k}
#now let's try a systematic method for setting k
kclusts <- tibble(k = 1:9) %>%
  mutate(kclust = map(k, ~kmeans(points, .x)),
         augmented = map(kclust, augment, points))
```

```{r assign}
#append cluster assignment to tibble
assignments <- kclusts %>%
  unnest(cols = c(augmented))
```

```{r plot_9_clust}
#Plot each model 
p1 <- ggplot(assignments, aes(x = x1, y = x2)) +
  geom_point(aes(color = .cluster), alpha = 0.8) +
  scale_color_brewer(palette = "Set3") +
  facet_wrap(~k)
p1
```

```{r elbow}
#Use a clustering function from {factoextra} to plot  total WSSs
fviz_nbclust(points, kmeans, "wss")
```


```{r more_fviz}
#Another plotting method
k3 <- kmeans(points, centers = 3, nstart = 25)

fviz_cluster(k3, geom = "point", data = points) + 
  ggtitle("k = 3")
```


In-class assignment!

Now it's your turn to partition a dataset.  For this round we'll use data from Roberts et al. 2008 on bio-contaminants in Sydney Australia's Port Jackson Bay.  The data are measurements of metal content in two types of co-occurring algae at 10 sample sites around the bay.

```{r data, message=FALSE}
#Read in data
metals_dat <- readr::read_csv(here::here("week-8/Harbour_metals.csv"))

# Inspect the data
head(metals_dat)

#Grab pollutant variables
metals_dat2 <- metals_dat[, 4:8] 
```

1. Start with k-means clustering - kmeans().  You can start with fviz_nbclust() to identify the best value of k. Then plot the model you obtain with the optimal value of k. 

```{r}
mod1 <- fviz_nbclust(metals_dat2, kmeans, "wss")

k3 <- kmeans(metals_dat2, centers = 3, nstart = 25)
fviz_cluster(k3, geom = "point", data = metals_dat2)
```


Do you notice anything different about the spacing between clusters?  Why might this be?

**- There is some overlap between the clusters. This might indicate there there are not actually 3 clusters and that what think is the optimal k from fviz_nbclust is not really the optimal k.** 


Run summary() on your model object.  Does anything stand out?

```{r}
mod1
```

**Looking at the mdodel, it seems like the elbow is at 3 clusters**

2. Good, now let's move to hierarchical clustering that we saw in lecture. The first step for that is to calculate a distance matrix on the data (using dist()). Euclidean is a good choice for the distance method.

```{r}
#get_dist()
dist_euc <- dist(metals_dat2, method = "euclidean")
```


2. Use tidy() on the distance matrix so you can see what is going on. What does each row in the resulting table represent?

```{r}
tidy(dist_euc)
```

**Each row in the table is the euclidean distance between the item1 and item2.**

3. Then apply hierarchical clustering with hclust().

```{r}
metal_clust <- hclust(dist_euc)
```

4. Now plot the clustering object. You can use something of the form plot(as.dendrogram()).  Or you can check out the cool visual options here: https://rpubs.com/gaston/dendrograms

```{r, fig.width=10}
plot(as.dendrogram(metal_clust))
```


How does the plot look? Do you see any outliers?  How can you tell?  

**The plot looks like a hierarchical structure and there seems to be 3 groups if the branches are cut at a cluster distance of around 75. Though there are potentially outliers on the far left (51-18) and seem to be the most different. These points take the longest to become nested with the rest of the dendrogram. If these are cut from the rest of the dendrogram, we would have two clusters.** 
