---
title: "Lab5"
author: "Sam Muir"
date: "2023-02-07"
output: html_document
---

This week's lab is a musical lab. You'll be requesting data from the Spotify API and using it to build k-nearest neighbor and decision tree models.

In order to use the Spotify API you must have a Spotify account. If you don't have one, sign up for a free one here: <https://www.spotify.com/us/signup>

Once you have an account, go to Spotify for developers (<https://developer.spotify.com/>) and log in. Click the green "Create a Client ID" button to fill out the form to create an app create an app so you can access the API.

On your developer dashboard page, click on the new app you just created. Go to Settings -\> Basic Information and you will find your Client ID . Click "View client secret" to access your secondary Client ID. Scroll down to Redirect URIs and enter: <http://localhost:1410/>

You have two options for completing this lab.

**Option 1**: **Classify by users**. Build models that predict whether a given song will be in your collection vs. a partner in class. This requires that you were already a Spotify user so you have enough data to work with. You will download your data from the Spotify API and then exchange with another member of class.

**Option 2**: **Classify by genres**. Build models that predict which genre a song belongs to. This will use a pre-existing Spotify dataset available from Kaggle.com (<https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify>)

```{r}
library(spotifyr) #API interaction
library(tidyverse)
library(tidymodels)
source("week-5/keys.R")
```

Client ID and Client Secret are required to create and access token that is required to interact with the API. You can set them as system values so we don't have to do provide them each time.

```{r access_API}
Sys.setenv(SPOTIFY_CLIENT_ID = SPOTIFY_CLIENT_ID) 
Sys.setenv(SPOTIFY_CLIENT_SECRET = SPOTIFY_CLIENT_SECRET)

authorization_code <- get_spotify_authorization_code(scope = scopes()[c(1:19)]) #sets an authorization code that you'll need to provide for certain get_ functions via my_tracks <- get_my_saved_tracks(authorization = authorization_code)

access_token <- get_spotify_access_token() #takes ID and SECRET, sends to Spotify and receives an access token
```

**Option 1: Data Preparation**

You can use get_my_saved_tracks() to request all your liked tracks. It would be good if you had at least 150-200 liked tracks so the model has enough data to work with. If you don't have enough liked tracks, you can instead use get_my_recently_played(), and in that case grab at least 500 recently played tracks if you can.

The Spotify API returns a dataframe of tracks and associated attributes. However, it will only return up to 50 (or 20) tracks at a time, so you will have to make multiple requests. Use a function to combine all your requests in one call.

```{r}
offsets = seq(from = 0, to = 150, by = 50)

#initializing an empty df 
my_tracks <- data.frame(matrix(nrow = 0, ncol = 30))

# function to get my 150 most recently liked tracks 
for (i in seq_along(offsets)) {  
  liked_tracks = get_my_saved_tracks(authorization = authorization_code, limit = 50, 
                                     offset = offsets[i])
  df_temp = as.data.frame(liked_tracks) # creating a temporary data frame to store the 50 liked tracks from a given run
  my_tracks <- rbind(my_tracks, df_temp) # binding the temporary data frame to my liked tracks data frame 
}
```

Once you have your tracks, familiarize yourself with this initial dataframe. You'll need to request some additional information for the analysis. If you give the API a list of track IDs using get_track_audio_features(), it will return an audio features dataframe of all the tracks and some attributes of them.

```{r}
audio1 <- get_track_audio_features(my_tracks$track.id[1:100])
audio2 <- get_track_audio_features(my_tracks$track.id[101:200])

audio_features <- rbind(audio1, audio2)
```


These track audio features are the predictors we are interested in, but this dataframe doesn't have the actual names of the tracks. Append the 'track.name' column from your favorite tracks database.

```{r}
sam_audio <- my_tracks %>%
  select(track.name) %>%
  bind_cols(audio_features) %>%
  mutate(name = "sam")
```


Find a class mate whose data you would like to use. Add your partner's data to your dataset. Create a new column that will contain the outcome variable that you will try to predict. This variable should contain two values that represent if the track came from your data set or your partner's.

```{r}
anna_audio <- read_csv("week-5/named_audio_anna.csv") %>%
  mutate(name = "anna")

both_audio <- rbind(sam_audio, anna_audio)
```


**Option 2: Data preparation**

Download the Spotify dataset from <https://www.kaggle.com/datasets/mrmorj/dataset-of-songs-in-spotify>

Inspect the data. Choose two genres you'd like to use for the classification task. Filter down the data to include only the tracks of that genre.

###Data Exploration (both options)

Let's take a look at your data. Do some exploratory summary stats and visualization.

For example: What are the most danceable tracks in your dataset? What are some differences in the data between users (Option 1) or genres (Option 2)?

```{r}
ggplot(both_audio, aes(x = danceability, fill = name)) +
  geom_density(alpha=0.5) +
  scale_fill_manual(values=c("magenta4", "seagreen"))+
  labs(x="Danceability", y="Density", title = "Distribution of Music Danceability Data") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal()


ggplot(both_audio, aes(x = energy, fill = name)) +
  geom_density(alpha=0.6) +
  scale_fill_manual(values=c("magenta4", "seagreen"))+
  labs(x="Energy", y="Density", title = "Distribution of Music Energy Data") +
  guides(fill=guide_legend(title="Listener"))+
  theme_minimal()
```


### **Modeling**

Create competing models that predict whether a track belongs to:

Option 1. you or your partner's collection

### KNN

```{r}
all_tracks_modeling <- both_audio %>%  
  mutate_if(is.ordered, .funs = factor, ordered = F) %>%
  select(-track.name, -type, -id, -uri, -track_href, -analysis_url) %>%
  mutate(name = as.factor(name))
```


```{r}
#splitting the data
set.seed(123)

#initial split of data
tracks_split <- initial_split(all_tracks_modeling, prop = 0.75)
tracks_train <- training(tracks_split)
tracks_test <- testing(tracks_split)
```

```{r}
# create a recipe
tracks_recipe <- recipe(name ~ ., data = tracks_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>% #normalize numeric to make sure scale is okay
  prep()
```

```{r}
# Define our KNN model with tuning
knn_spec_tune <- nearest_neighbor(neighbor = tune()) %>%
  set_mode("classification") %>%
  set_engine("kknn")

# Check the model
knn_spec_tune
```

```{r}
# Define a new workflow
wf_knn_tune <- workflow() %>%
  add_model(knn_spec_tune) %>%
  add_recipe(tracks_recipe)

# 10-fold CV on the training dataset
set.seed(123)

cv_folds <- tracks_train %>%
  vfold_cv(v = 10)


# Fit the workflow on our predefined folds and hyperparameters
fit_knn_cv <- wf_knn_tune %>%
  tune_grid(resamples = cv_folds,
            grid = 10)
    
# Check the performance with collect_metrics()
collect_metrics(fit_knn_cv)

autoplot(fit_knn_cv) + # plot cv results for parameter tuning
  theme_bw()
```


```{r}
# The final workflow for our KNN model
final_wf <- wf_knn_tune %>%
  finalize_workflow(select_best(fit_knn_cv, metric = "roc_auc"))

train_fit = fit(final_wf, tracks_train)

train_predict = predict(object = train_fit, new_data = tracks_train) %>% #predict the training set
  bind_cols(tracks_train) #bind training set column to prediction

test_predict = predict(train_fit, tracks_test) %>% #get prediction probabilities for test 
  bind_cols(tracks_test) %>%  #bind to testing column
  mutate(name = as.factor(name))
```


```{r}
accuracy(train_predict, truth = name, estimate = .pred_class) #get training accuracy

sensitivity(train_predict, truth = name, estimate = .pred_class)
specificity(train_predict, truth = name, estimate = .pred_class)
```

```{r}
accuracy(test_predict, truth = name, estimate = .pred_class) #get accuracy of testing prediction

sensitivity(test_predict, truth = name, estimate = .pred_class)
specificity(test_predict, truth = name, estimate = .pred_class)
```


```{r}
# Optimizing workflow
final_fit <- final_wf %>% 
  last_fit(tracks_split)

# Collect metrics on the test data!
final_fit %>% 
  collect_metrics()
```

```{r}
test_predict %>%
  conf_mat(truth = name, estimate = .pred_class) %>%
  autoplot(type = 'heatmap') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```


## Decision Tree

```{r}
# Modeling packages
library(rpart)       # direct engine for decision tree application
library(caret)       # meta engine for decision tree application

# Model interpretability packages
library(rpart.plot)  # for plotting decision trees
library(vip)         # for feature importance
library(pdp)
```

```{r}
#Preprocess the data
tracks_rec <- recipe(name ~ ., data = tracks_train) %>%
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%
  step_normalize(all_numeric_predictors())
```

```{r}
#new spec, tell the model that we are tuning hyperparams
tree_spec_tune <- decision_tree(
  cost_complexity = tune(),
  tree_depth = tune(),
  min_n = tune()
) %>%
  set_engine("rpart") %>%
  set_mode("classification")

tree_grid <- grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)
tree_grid
```

```{r workflow_tree}
wf_tree_tune <- workflow() %>%
  add_recipe(tracks_recipe) %>%
  add_model(tree_spec_tune)
```

```{r resampling}
#set up k-fold cv. This can be used for all the algorithms
tracks_cv <- tracks_train %>%
  vfold_cv(v=10)

tracks_cv
```

```{r}
doParallel::registerDoParallel() #build trees in parallel
#200s
system.time(
  tree_rs <- tune_grid(
    wf_tree_tune,
    resamples = tracks_cv,
    grid = tree_grid,
    metrics = metric_set(accuracy)
  )
)

tree_rs
```

Use autoplot() to examine how different parameter configurations relate to accuracy 
```{r}
autoplot(tree_rs) +
  theme_bw()
```

```{r select_hyperparam}
show_best(tree_rs) # showing
select_best(tree_rs) # what we'll input
```

We can finalize the model specification where we have replaced the tune functions with optimized values.

```{r final_tree_spec}
final_tree <- finalize_workflow(wf_tree_tune, select_best(tree_rs))

final_tree
```

This model has not been fit yet though.

```{r final_tree_fit}
#similar functions here.
final_tree_fit <- fit(final_tree, data = tracks_train) 

# last_fit() fit on the training data but then also evaluates on the testing data
final_tree_result <- last_fit(final_tree, tracks_split)

# testing predictions and metrics
final_tree_result$.predictions 
predict_data <- as.data.frame(final_tree_result$.predictions) %>%
  bind_cols(tracks_test)

final_tree_result$.metrics
```

#Visualize variable importance
```{r tree_vip}
final_tree_fit %>%
  vip(geom = "point") +
  theme_bw()

# differences between groups
ggplot(predict_data, aes(.pred_class, loudness)) +
  geom_boxplot()
```


Option 2. genre 1 or genre 2

You will eventually create four final candidate models:

1.  k-nearest neighbor (Week 5)
2.  decision tree (Week 5)
3.  bagged tree (Week 6)
    -   bag_tree()
    -   Use the "times =" argument when setting the engine during model specification to specify the number of trees. The rule of thumb is that 50-500 trees is usually sufficient. The bottom of that range should be sufficient here.
4.  random forest (Week 6)
    -   rand_forest()
    -   m_try() is the new hyperparameter of interest for this type of model. Make sure to include it in your tuning process

Go through the modeling process for each model:

Preprocessing. You can use the same recipe for all the models you create.

Resampling. Make sure to use appropriate resampling to select the best version created by each algorithm.

Tuning. Find the best values for each hyperparameter (within a reasonable range).

Compare the performance of the four final models you have created.

Use appropriate performance evaluation metric(s) for this classification task. A table would be a good way to display your comparison. Use at least one visualization illustrating your model results.
