---
title: "Lab 8"
author: "Mateo Robbins"
date: "2024-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(dplyr)
library(kernlab)
library(tidyverse)
```

## Forest Cover Classification with SVM

In this week's lab we are exploring the use of Support Vector Machines for multi-class classification. Specifically, you will be using cartographic variables to predict forest cover type (7 types).

Natural resource managers responsible for developing ecosystem management strategies require basic descriptive information including inventory data like forest cover type for forested lands to support their decision-making processes. However, managers generally do not have this type of data for in-holdings or neighboring lands that are outside their immediate jurisdiction. One method of obtaining this information is through the use of predictive models.

You task is build both an SVM and a random forest model and compare their performance on accuracy and computation time.

1.  The data is available here: <https://ucsb.box.com/s/ai5ost029enlguqyyn04bnlfaqmp8kn4>

Explore the data.

```{r}
forest <- read_csv("week-9/covtype_sample.csv") %>%
  janitor::clean_names() %>%
  mutate(cover_type = as.factor(cover_type)) %>%
  mutate(across(6:54, as.factor)) %>%
  select(-soil_type_15, -soil_type_37, -soil_type_38, -soil_type_39, -soil_type_40) # only have value 

skimr::skim(forest)
```

```{r}
ggplot(forest, aes(x = cover_type)) +
  geom_histogram(stat = "count") +
  theme_bw()
```

-   What kinds of features are we working with?

**We have some topographical features that are continuous variables (slope, aspect) and some that are binary (the soil types).**

-   Does anything stand out that will affect you modeling choices?

**Some of the continuous variables are not normally distributed. There are also both continuous and categorical variables.**

Hint: Pay special attention to the distribution of the outcome variable across the classes.

2.  Create the recipe and carry out any necessary preprocessing. Can you use the same recipe for both models?

3.  Create the folds for cross-validation.

4.  Tune the models. Choose appropriate parameters and grids. If the computational costs of tuning given your strategy are prohibitive, how might you work around this?

5.  Conduct final predictions for both models and compare their prediction performances and computation costs from part 4.

```{r}
# Split the data into training and testing
split <- initial_split(forest, 0.7, strata = cover_type)
forest_training <- training(split)
forest_testing <- testing(split)

# create recipe
svm_rec <- recipe(cover_type~., data = forest_training) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors())

svm_linear_spec <- svm_poly(degree = 1,
                            cost = tune()) %>% # cost parameter: width of margin around hyper plane
  set_mode("classification") %>%
  set_engine("kernlab")

svm_linear_wf <- workflow() %>% 
  add_model(svm_linear_spec %>% 
              set_args(cost_tune())) %>% 
  add_formula(cover_type~.)
svm_linear_wf

set.seed(258)
forest_fold <- vfold_cv(forest_training, strata = cover_type, v = 5)

doParallel::registerDoParallel(cores = 4) 

param_grid <- grid_regular(cost(), levels = 5)

# tune grid
system.time(
tune_res <- tune_grid(
  svm_linear_wf,
  resamples = forest_fold,
  grid = param_grid
))
#      user    system   elapsed 
# 19934.777  1483.132  6402.373 

#write_rds(tune_res, "week-9/svm_tune.rda")
tune_res <- read_rds("week-9/svm_tune.rda")

autoplot(tune_res)

## finalize 
best_cost <- select_best(tune_res, metric = "accuracy")
best_cost

svm_linear_final <- finalize_workflow(svm_linear_wf, best_cost)

svm_linear_fit <- svm_linear_final %>% fit(forest_training)
#write_rds(svm_linear_fit, "week-9/svm_linear_fit.rda")
svm_linear_fit <- read_rds("week-9/svm_linear_fit.rda")
svm_linear_fit

# augment 
svm_linear_augment <- augment(svm_linear_fit, new_data = forest_testing) 
#write_rds(svm_linear_augment, "week-9/svm_linear_augment.rda")
svm_linear_augment <- read_rds("week-9/svm_linear_augment.rda")

svm_linear_augment %>%
  conf_mat(truth = cover_type, estimate = .pred_class) %>%
  autoplot(type = "heatmap")

svm_linear_augment %>%
  roc_curve(truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7) %>%
  autoplot()

svm_linear_augment %>%
  roc_auc(truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7)
```

```{r}
#svm rbf --------------------
svm_rbf_spec <- svm_rbf() %>% # cost parameter: width of margin around hyper plane
  set_mode("classification") %>%
  set_engine("kernlab")

system.time(
svm_rbf_fit <- svm_rbf_spec %>%
  fit(cover_type~., data = forest_training))
#    user  system elapsed 
# 299.652  14.191 315.562 
#write_rds(svm_rbf_fit, "week-9/svm_rbf_fit.rda")
svm_rbf_fit <- read_rds("week-9/svm_rbf_fit.rda")
svm_rbf_fit

#Examine model performance via confusion matrix
svm_rbf_augment <- augment(svm_rbf_fit, new_data = forest_testing)
#write_rds(svm_rbf_augment, "week-9/svm_rbf_augment.rda")
svm_rbf_augment <- read_rds("week-9/svm_rbf_augment.rda")

svm_rbf_augment %>%
  conf_mat(truth = cover_type, estimate = .pred_class) %>%
  autoplot(type = "heatmap")

#We can examine our model's performance using ROC and AUC
svm_rbf_augment %>%
  roc_curve(., truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7) %>%
  autoplot()

svm_rbf_augment %>%
  roc_auc(truth = cover_type, .pred_1, .pred_2, .pred_3, .pred_4, .pred_5, .pred_6, .pred_7)
```

```{r}
# random forest ----
second_split <- initial_split(forest_training, 0.05)
second_training <- training(second_split)
# create a recipe
rf_recipe <- recipe(cover_type ~ ., data = second_training) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>% #normalize numeric to make sure scale is okay
  prep()

rf_model <- rand_forest(mtry = tune(),
                  trees = tune()) %>%
  set_engine("ranger") %>%
  set_mode("classification")

# workflow ----
rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

# parameter tuning ----
system.time(
  rf_cv_tune <- rf_workflow %>%
  tune_grid(resamples = forest_fold, grid = 2)) #use cross validation to tune mtry and trees parameters

#get metrics from tuning cv to pick best model ----
#collect_metrics(rf_cv_tune) 

#plot cv results for parameter tuning ----
autoplot(rf_cv_tune) + 
  theme_bw()

# finalize workflow ----
rf_best <- show_best(rf_cv_tune, n = 1, metric = "roc_auc") #get metrics for best random forest model
rf_best

rf_final <- finalize_workflow(rf_workflow,
                              select_best(rf_cv_tune, metric = "roc_auc"))

# model fitting ----
train_fit_rf <- fit(rf_final, tracks_train) #fit the KNN model to the training set
#train_fit_rf

# prediction probabilities ----
test_predict_rf <- predict(train_fit_rf, tracks_test) %>% #get prediction probabilities for test 
  bind_cols(tracks_test) %>%  #bind to testing column
  mutate(name = as.factor(name))

test_predict2_rf <- predict(train_fit_rf, tracks_test, type = "prob") %>% #get testing prediction
  bind_cols(tracks_test) %>%  #bind to testing column
  mutate(name = as.factor(name))

# model metrics and evaluation
#accuracy(test_predict_rf, truth = name, estimate = .pred_class) #get accuracy of testing prediction

rf_final_result <- last_fit(rf_final, tracks_split)

# testing predictions and metrics ----
#rf_final_result$.predictions 
rf_predict_data <- as.data.frame(rf_final_result$.predictions) %>%
  bind_cols(tracks_test)

rf_final_result$.metrics

# roc auc curve
test_roc_auc_rf <- roc_curve(test_predict2_rf, name, .pred_anna)

rf_metrics <- rf_final_result$.metrics[[1]]

```

-   Which type of model do you think is better for this task?
-   Why do you speculate this is the case?
