---
title: "Lab6"
author: "Sam Muir"
date: "2023-03-01"
output: html_document
---

## Case Study: Eel Distribution Modeling

This week's lab follows a project modeling the eel species Anguilla australis described by Elith et al. (2008). There are two data sets for this lab.  You'll use one for training and evaluating your model, and you'll use your model to make predictions predictions on the other.  Then you'll compare your model's performance to the model used by Elith et al.

## Data

Grab the training data sets (eel.model.data.csv, eel.eval.data.csv) from github here:
https://github.com/MaRo406/eds-232-machine-learning/blob/main/data 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(skimr)
library(tidymodels)
library(caret)
library(patchwork)
library(rsample)
```

```{r}
eel_model_data <- read_csv("week-7/eel.model.data.csv") %>%
  mutate(Angaus = as.factor(Angaus)) %>%
  select(-Site)
eel_eval_data <- read_csv("week-7/eel.eval.data.csv")
```

### Split and Resample

Split the model data (eel.model.data.csv) into a training and test set, stratified by outcome score (Angaus). Use 10-fold CV to resample the training set.

```{r}
set.seed(222)
eel_split <- initial_split(eel_model_data, strata = Angaus)
eel_training <- training(eel_split)
eel_testing <- testing(eel_split)

#set up k-fold cv
eel_cv <- eel_training %>%
  vfold_cv(v=10)
eel_cv
```


### Preprocess

Create a recipe to prepare your data for the XGBoost model

```{r}
eel_recipe <- recipe(Angaus ~ ., data = eel_training) %>% #create model recipe
  step_dummy(all_nominal_predictors()) %>% #create dummy variables from all factors
  step_normalize(all_numeric_predictors()) #normalize all numeric predictors
```


## Tuning XGBoost

### Tune Learning Rate

Following the XGBoost tuning strategy outlined in lecture, first we conduct tuning on just the learning rate parameter:

1.  Create a model specification using {xgboost} for the estimation

-   Only specify one parameter to tune()

```{r}
eel_model_learn <- boost_tree(learn_rate = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

eel_workflow_learn <- workflow() %>% #create workflow
  add_model(eel_model_learn) %>%
  add_recipe(eel_recipe)
```


2.  Set up a grid to tune your model by using a range of learning rate parameter values: expand.grid(learn_rate = seq(0.0001, 0.3, length.out = 30))

-   Use appropriate metrics argument(s) - Computational efficiency becomes a factor as models get more complex and data get larger. Record the time it takes to run. Do this for each tuning phase you run.You could use {tictoc} or Sys.time().

```{r}
grid <- expand.grid(learn_rate = seq(0.0001, 0.3, length.out = 30))

system.time(eel_cv_tune <- eel_workflow_learn %>%
  tune_grid(resamples = eel_cv, grid = grid))
```


3.  Show the performance of the best models and the estimates for the learning rate parameter values associated with each.

```{r}
collect_metrics(eel_cv_tune)
```


### Tune Tree Parameters

1.  Create a new specification where you set the learning rate (which you already optimized) and tune the tree parameters.

```{r}
eel_model <- boost_tree(learn_rate = tune(), trees = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

eel_workflow <- workflow() %>% #create workflow
  add_model(eel_model) %>%
  add_recipe(eel_recipe)
```


2.  Set up a tuning grid. This time use grid_latin_hypercube() to get a representative sampling of the parameter space

```{r}
grid_hypercube <- grid_latin_hypercube(learn_rate(), trees())
  
system.time(eel_full_tune <- eel_workflow %>%
  tune_grid(resamples = eel_cv, grid = grid_hypercube))
```


3.  Show the performance of the best models and the estimates for the tree parameter values associated with each.

```{r}
collect_metrics(eel_full_tune)
```


### Tune Stochastic Parameters

1.  Create a new specification where you set the learning rate and tree parameters (which you already optimized) and tune the stochastic parameters.

```{r}
eel_model_3 <- boost_tree(learn_rate = tune(), trees = tune(), min_n = tune(), tree_depth = tune(), stop_iter = tune(), loss_reduction = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

eel_workflow_3 <- workflow() %>% #create workflow
  add_model(eel_model_3) %>%
  add_recipe(eel_recipe)
```


2.  Set up a tuning grid. Use grid_latin_hypercube() again.

```{r}
grid_hypercube <- grid_latin_hypercube(learn_rate(), trees(), min_n(), tree_depth(), stop_iter(), loss_reduction())
  
system.time(eel_tune_3 <- eel_workflow_3 %>%
  tune_grid(resamples = eel_cv, grid = grid_hypercube))
```


3.  Show the performance of the best models and the estimates for the tree parameter values associated with each.

```{r}
collect_metrics(eel_tune_3)
```


## Finalize workflow and make final prediction

1.  How well did your model perform? What types of errors did it make?

```{r}
eel_best = show_best(eel_tune_3, n = 1, metric = "roc_auc") #get metrics for best random forest model
eel_best

eel_final <- finalize_workflow(eel_workflow,
                              select_best(eel_tune_3, metric = "roc_auc"))
```


## Fit your model the evaluation data and compare performance

1.  Now used your final model to predict on the other dataset (eval.data.csv)

```{r}
train_fit_eel = fit(eel_final, eel_training)
train_fit_eel

test_predict_eel = predict(train_fit_eel, eel_eval_data) %>% #get prediction probabilities for test 
  bind_cols(eel_eval_data) %>%
  mutate(Angaus_obs = as.factor(Angaus_obs))
```


2.  How does your model perform on this data?

```{r}
accuracy(test_predict_eel, truth = Angaus_obs, estimate = .pred_class) #get accuracy of testing prediction
```


3.  How do your results compare to those of Elith et al.?

-   Use {vip} to compare variable importance

```{r}
train_fit_eel %>%
  vip::vip(geom = "point") +
  theme_bw()
```

**For both my results and the results of Elith et al., SegSumT has the highest variable importance and USNative has the second highest.**

-   What do your variable importance results tell you about the distribution of this eel species?

**The variable SegSumT tells us that the Summer air temperature has the biggest effect on Angaus occurence.**
